Distributed Web Browser Mining for Ethereum

### 2018-03-21 paper/paper.tex

\subsection{Ethash}
Ethereum uses a Proof-of-Work algorithm called Ethash for verification of blocks. Derived from the Dagger and Hashimoto algorithms ~\cite{dagger-hashimoto}, Ethash operates by first generating a 256 bit seed. This seed is different for every epoch (30,000 blocks). For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash. This seed is then used to compute a pseudo-random cache. The cache production process involves using the seed hash to first sequentially filling up X bytes of memory (size of the cache), then performing Y passes of the RandMemoHash algorithm ~\cite{randmemohash}. This transforms the initial cache to give us the final value for the cache. This cache is simply a byte array of 64-byte values ~\cite{Ethmining}. Lightweight clients may use this cache for verification, while full node clients can use it to calculate a DAG (Directed Acyclic Graph) file. Essentially, each node in the DAG is generated by combining data from 256 pseudorandomly selected cache nodes and hashing them together. This process is repeated to generate all nodes in the DAG. Mining is performed by starting with the current block header hash and nonce, which are combined to get a 128B wide “Mix” (another byte array). The Mix is then used to fetch a specific slice of the DAG from memory. (Each slice is 512 bits long.) After this, the Mix is updated with the fetched part of the DAG. Then, this updated Mix is used to fetch a new part of the DAG. This process of sequentially fetching parts of the DAG is repeated 64 times, and the final value of the Mix is put through a transformation to obtain a 32 byte digest. This digest is then compared to the threshold, and if it is lesser than it, the nonce is valid and the block is successfully mined and can be broadcast to the network. However, if the digest is greater than the threshold, the nonce was unsuccessful, and the entire process must be repeated with a new nonce ~\cite{Ethmining}. This procedure for mining is depicted in figure ~\ref{fig:ethash}. It is important to note that the parts of the DAG that are used to compute the hash for a particular block depend on the nonce used, hence no way to pre-determine which parts will be useful to have in memory. Hence this forces miners to store entire DAG, making mining "Memory Hard". This is intended to even the playing field, as ASIC miners no longer have any advantage as they do for Bitcoin. GPU miners, however, do have an advantage, normally having hashrates of over 10 times that of a CPU miner.
While mining is memory intensive, verification is relatively lightweight. This is because of the property that each ``item'' in the DAG depends on a set of pseudo-randomly selected items from the cache. Hence, the cache is used to regenerate only the specific pieces of the DAG that are needed to recalculate the hash for the particular nonce. And so only the cache needs to be stored by clients that just perform verification.

@@

\subsection{Information theoretic bounds}
With this lazy-evaluation based approach, we can calculate exactly how many hashes we would have to compute in order to fill up a buffer of a given size. This approach is a variation of the coupon collector problem \cite{couponCollector}. For our purposes, let us assume the following:
\begin{gather}
    Number\;of\;DAG\;slices = n\\
    Number\;slices\;buffer\;can\;store = i
\end{gather}
If we have buffered \verb|x| slices so far, the probability of obtaining a new slices is given by:
\begin{gather}
  P_x = \frac{i-x}{n}
\end{gather}
The event of obtaining a new slice after having already collected \verb|x| out of the \verb|i| slices that can be buffered can be modeled as a geometric distribution. Hence, the expected number of trials $t_x$ it takes to get a new  slice after having seen \verb|x| out of the \verb|i| slices is:
\begin{gather}
  E(t_x) = \frac{1}{P_x} = \frac{n}{i-x}
\end{gather}
By linearity of expectation, the total number of trials it would take to see all \verb|x| slices is:
\begin{gather}
  E(t) = \sum_{x=0}^{i-1}E(t_x)\\
    E(t) = \sum_{x=0}^{i-1}P_x = \sum_{i=0}^{i-1}\frac{n}{i-x} = n\sum_{i=0}^{i-1}\frac{1}{i-x}
\end{gather}
Since the summation represents the harmonic series, the expected value can be approximated as follows:
\begin{gather}
  E(t) = n\log{i}
\end{gather}
Now, let us do a simple calculation for one extreme case, where \verb|i| = \verb|n| (i.e., we allocate a buffer as large as the entire DAG on the browser), in order to get an upper bound on the number of hashes it would take to fill this buffer up. The private test network for Ethereum we used had a total of 16777186 slices in the DAG, which means it would take:
\begin{gather}
  E(t) = 16777186*\log{16777186} = 2.79*10^8\;trials
\end{gather}
Now, each hash involved 64 random accesses of the DAG, which means the total number of hashes it would take to buffer the entire DAG is:
\begin{gather}
  Hashes = \frac{2.79*10^8}{64} = 4.36*10^6
\end{gather}
Hence, it takes around 4 million hashes to fill up a buffer as large as the entire DAG. If $i < n$, the number of hashes should be less than 4 million. Since 4 million hashes is about as long as it could take to fill up the buffer, all of our experiments involve computing 4 million hashes, as we will see below.
Furthermore, let:
\begin{gather}
  Buffer\;hit\;rate = R \\
    Number\;of\;hashes\;computed = H
\end{gather}
Then,
\begin{gather}
  \lim_{H\to\infty}R = \frac{i}{n}
\end{gather}
This is because as $H\to\infty$, the buffer becomes full. Which means for any random slice of the DAG we pull out, the probability of it being in the buffer is $\frac{i}{n}$.

@@

\begin{figure}[H]
\centering
\includegraphics[width=250px,keepaspectratio]{Hybrid-Miner.pdf}
\caption{\label{fig:hybridArchitecture} JavaScript Cache Miner Architecture}
\end{figure}

>>>>

\subsection{Ethash}
Ethereum uses a Proof-of-Work algorithm called Ethash for mining blocks. Derived from the Dagger and Hashimoto algorithms ~\cite{dagger-hashimoto}, Ethash operates by first generating a 256 bit seed. This seed is different for every epoch (30,000 blocks). For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros. For every other epoch it is always the Keccak-256 hash of the previous seed hash. This seed is then used to compute a pseudo-random cache. The cache production process involves using the seed hash to first sequentially filling up X bytes of memory (size of the cache), then performing Y passes of the RandMemoHash algorithm ~\cite{randmemohash}. This transforms the initial cache to give us the final value for the cache. This cache is simply an array 4 byte integers ~\cite{Ethmining}. Lightweight clients may use this cache for verification, while full node clients can use it to calculate a DAG (Directed Acyclic Graph) file. Essentially, each node in the DAG is generated by combining data from 256 pseudorandomly selected cache nodes and hashing them together. This process is repeated to generate all nodes in the DAG. Mining is performed by starting with the current block header hash and nonce, which are combined to get a 128B wide ``Mix'' (another byte array). The Mix is then used to fetch a specific slice of the DAG from memory. (Each slice is 512 bits long, or 16 4 byte integers) After this, the Mix is updated with the fetched part of the DAG. Then, this updated Mix is used to fetch a new part of the DAG. This process of sequentially fetching parts of the DAG is repeated 64 times, and the final value of the Mix is put through a transformation to obtain a 32 byte digest. This digest is then compared to the threshold, and if it is lesser than it, the nonce is valid and the block is successfully mined and can be broadcast to the network. However, if the digest is greater than the threshold, the nonce was unsuccessful, and the entire process must be repeated with a new nonce ~\cite{Ethmining}. This procedure for mining is depicted in figure ~\ref{fig:ethash}. It is important to note that the parts of the DAG that are used to compute the hash for a particular block depend on the nonce used, hence no way to pre-determine which parts will be useful to have in memory. Hence this forces miners to store entire DAG in memory, making mining ``Memory Hard''. This is intended to even the playing field, as ASIC miners no longer have any advantage as they do for Bitcoin. GPU miners, however, do have an advantage, normally having hashrates of over 10 times that of a CPU miner.
While mining is memory intensive, verification is relatively lightweight. This is because of the property that each ``slice'' in the DAG depends on a set of pseudo-randomly selected items from the Cache. Hence, the Cache is used to regenerate only the specific slices of the DAG that are needed to recalculate the hash for the particular nonce. And so only the Cache needs to be stored by clients that just perform verification. We use this property of being able to generate the parts of the DAG as needed to our advantage in order to alleviate some of the memory and network bandwidth restrictions that browsers typically face.

@@

\subsection{Information theoretic bounds}
With this lazy-evaluation based approach, we can calculate exactly how many hashes we would have to compute in order to fill up a buffer of a given size. This approach is a variation of the coupon collector problem \cite{couponCollector}. For our purposes, let us assume the following:
\begin{gather}
    Number\;of\;DAG\;slices = n\\
    Size\;of\;buffer\;as\;a\;fraction\;of\;the\;total DAG = i\;;i=[0,1]
\end{gather}
If we have buffered \verb|x| slices so far, the probability of obtaining a new slice is given by:
\begin{gather}
  P_x = \frac{ni-x}{n}
\end{gather}
The event of obtaining a new slice after having already collected \verb|x| out of the \verb|ni| slices that can be buffered can be modeled as a geometric distribution. Hence, the expected number of trials $t_x$ it takes to get a new  slice after having seen \verb|x| out of the \verb|i| slices is:
\begin{gather}
  E(t_x) = \frac{1}{P_x} = \frac{n}{ni-x}
\end{gather}
By linearity of expectation, the total number of trials it would take to see all \verb|x| slices is:
\begin{gather}
  E(t) = \sum_{x=0}^{ni-1}E(t_x)\\
    E(t) = \sum_{x=0}^{ni-1}P_x = \sum_{x=0}^{ni-1}\frac{n}{ni-x} = n\sum_{x=0}^{ni-1}\frac{1}{ni-x}
\end{gather}
Since the summation represents the harmonic series, the expected value can be approximated as follows:
\begin{gather}
  E(t) = n\log{ni}
\end{gather}
Now, equation ~\ref{eq:7} is in terms of the number of individual accesses of the DAG. In order to get the expected number of Hashes (as opposed to the expected number of individual DAG slice accesses), we need to divide by 64 since each hash involves 64 accesses of the DAG. Hence if we define: 
\begin{gather}
    Number\;of\;hashes\;computed = H
\end{gather}
\begin{gather}
  E(H) = \frac{n\log{ni}}{64}
\end{gather}
Now, let us do a simple calculation for one extreme case, where $i = 1$ (i.e., we allocate a buffer as large as the entire DAG on the browser), in order to get an upper bound on the number of hashes it would take to fill this buffer up. The private test network for Ethereum we used had a total of 16777186 slices in the DAG, which means it would take:
\begin{gather}
  E(H) = \frac{16777186*\log{16777186}}{64} = 4.36*10^6\;hashes
\end{gather}
Hence, it takes around 4 million hashes to fill up a buffer as large as the entire DAG. If $i < n$, the number of hashes should be less than 4 million. So effectively, 4 million hashes is a very lose upper bound on the number of hashes needed to fill up a DAG buffer.

Now let us say that we do not want to wait for the entire buffer to fill up -- how many hashes would we have to compute until a reasonable fraction of the buffer is full? Concretely speaking, let us assume the following:
\begin{gather}
  Fraction\;of\;DAG\;buffer\;to\;be\;filled = f;\;f \in [0,1]
\end{gather}
Then, for a buffer with space for \verb|i| slices, the number of slices we want to fill it with in order for a fraction \verb|f| of it to be full would be: 
\begin{gather}
  Number\;of\;slices\;to\;fill\;DAG\;buffer\;with = nfi
\end{gather}
This would modify equation \ref{eq:6} for the expected number of trials to fill up a fraction \verb|f| of the DAG buffer to be as follows:
\begin{gather}
    E(t) = \sum_{x=0}^{nfi}P_x = \sum_{i=0}^{nfi}\frac{n}{ni-x} = n\sum_{x=0}^{nfi}\frac{1}{ni-x}\\
    E(t) = n\sum_{x=0}^{nfi}\frac{1}{ni-x}
\end{gather}
This is essentially an early termination of the summation, and can be expressed as follows:
\begin{gather}
    E(t) = n\sum_{i=0}^{nfi}\frac{1}{ni-x} = n(\sum_{x=0}^{ni-1}\frac{1}{ni-x} - \sum_{x=nfi}^{ni-1}\frac{1}{ni-x})\\
    E(t) = n(\log{ni} - \log{(ni-nfi)}) \\
    E(t) = n\log{\frac{1}{1-f}}
\end{gather}
Again, we divide by 64 to obtain the expected number of hash computations:
\begin{gather}
    E(H) = \frac{n\log{\frac{1}{1-f}}}{64}
\end{gather}
Interestingly \verb|E(H)| does not depend on \verb|i|; the size of the DAG buffer as a fraction of the entire DAG. Figure X shows a graph of how \verb|E(H)| varies as a function of the fraction of the buffer that we want to be full, \verb|f| if we set $n = 16777186$ (the total number of DAG slices on the experimental private test network).

From figure X we see that even computing merely 800,000 hashes fills our DAG buffer by more than 95\%! Beyond that, the marginal cost of computing more hashes outweighs the 
Finally, let us compute what the buffer hit rate should be for a given buffer size when the number of hashes computed tends to infinity. For this, let:
\begin{gather}
  Buffer\;hit\;rate = R \\
    Number\;of\;hashes\;computed = H
\end{gather}
Then,
\begin{gather}
  \lim_{H\to\infty}R = \frac{i}{n}
\end{gather}
This is because after a finite number of hashes, the buffer becomes full. Which means for any random slice of the DAG we pull out, the probability of it being in the buffer is $\frac{i}{n}$. However, the buffer hit rates were extremely low when H was close to 0 low $H\to\infty$

@@

\begin{figure}[H]
\centering
\includegraphics[width=250px,keepaspectratio]{Hybrid-Miner.pdf}
\caption{\label{fig:hybridArchitecture} Web Miner Architecture}
\end{figure}
